This dissertation is a product made by me, but I am a product made by many others.
I've benefitted greatly from having the following people in my life.

My family, especially my parents Minh and Phiet, my sister Tracie and her husband Jeremy,
my uncle Le and aunt Thuy, my sisters Mai and Ann, my brother-in-law Nash,
and my nieces Nadia and Maria.

I also wish to thank the following people and things in my life in no particular order.
You all supported me, gave me ideas, and helped me grow.

My coworking group, fellow tEps, and fastidious last-minute proofreaders: Blake Stacey and Eric Downes.

The quantum theory group at UW, and then MIT: Kamil Michnicki, Lukas Svec, Kevin Zatloukal, Isaac Crosson, David Rosenbaum;

My housemates at 419 Boylston: Mike Toomim, Nate Morris, and Marius Nita.

The grad students I started out with: Nodira Khoussainova, Kayur Patel, Liz Tseng, Travis Kriplean, Yaw Anokwa, Suporn Pongnumkul, Anna Gringauze, and Jonah Cohen.

My brother from another mother and graduation buddy: Dr. Alex B. Jaffe, esq.

My friends and fellow UW grad students: Eva Ringstrom, Ian Simon, Widad Machmouchi, Adrian Sampson, Mark Yatskar, Summer Thyme, Cyrus Rashtchian, Thach Nguyen, and Nguyen Nguyen.

My workout buddies Ray Cheng, Dimi Gklezakos, and Yun-En Liu.

These fine UW professors: Gaetano Borriello, David Notkin, Oren Etzioni, Steve Tanimoto.

Mike Ernst, who has been my mentor and role model at both MIT and UW.

The Evergreen State College, and everyone else who has ever given me a job.

My officemates: Jinna Lei, Ezgi Mercan, Kivanc Muslu, Gabe Schubiner,
Nicholas Fitzgerald, Bryan Ferris, Stanley Kok, Masaharu Kobashi,
Punya Biswal.

My number one cheerleader and daughter of the Pacific Northwest: Maggie Stringfellow Herring.

My friends at MIT for both the first and second times:
Adrian Bischoff, Jason Rolfe,
Malima Wolf, Brian Neltner, Keith Durand, Mark Tobenkin, Dan Taub, and
Josh Gordonson.

My Seattle friends: Abbie Smith-Howe, Justin Nafziger, Nick Braddy, Corinne Pascale, and Kate Bowlus.

My students at UW, especially Noah Siegel, Andrea McCool, Harshad Petwe, and Jeffrey Booth.

My design classmates at UW, especially Katherine Wong and Claire Simon.

My Burning Man campmates: Eric Miles, Lee Granas, David Morris, Ivan, Julianna, and Igor Mordatch.

My extended Burning Man family: Kether Axelrod, Tibet Boyer, Robert Jaffe, Taliesin Jaffe, and Anika Levy.

My life coach, Bi Lai.

These fine entertainments: Futurama, the tenth season; Star Treks, both The Next Generation and Deep Space Nine; Baz Luhrmann's Great Gatsby 3D movie and soundtrack; Stephen Colbert and Jon Stewart;
Florence + the Machine; The National; Ellie Goulding; Imagine Dragons.

My thesis committee: Paul Beame, Mark Oskin, and Boris Blinov.

My Microsoft Research mentor Krysta.

And finally I would like to thank my advisor Aram,
who is always willing to listen and to teach.

This work was supported by
the IARPA QCS program (D11PC20167).  The
views and conclusions contained herein are those of the authors and
should not be interpreted as necessarily representing the official
policies or endorsements, either expressed or implied, of IARPA,
DoI/NBC, or the U.S. Government.
The work in Chapter \ref{chap:factor-polylog}
was partially supported by Microsoft Research.