\chapter{Introduction}
\label{chap:intro}

% I. Quantum computing and perspectives
The combination of existing scientific disciplines provides fertile ground
for the development of entirely new scientific discoveries. Computer science
and quantum physics are now coming together in this way to give birth to
quantum computing. Computer science and quantum physics originated
as contemporaries in the early part of the 20th century, culminating in
technology that decisively ended World War II. They have remained
largely independent until recently, but the common thread connecting them
has always existed.

From the perspective of computer science, engineering is problem-driven.
We humans wish to solve a problem, like calculating ballistic weapon
trajectories, and we work backwards to electromechanical relays or punched
cards. To calculate,
we need to perform fast
arithmetic on stored numbers, for which we need to design devices in a
certain configuration, for which we need certain materials from nature.
At heart, it is the study of structured problem-solving which historically
has a strong
mechanical bias. This perspective asks the questions: what do we want to do,
and how can we
do it? It is like a child going into a LEGO store with a pre-existing plan:
the child only buys the bricks needed, or orders custom bricks.
It is an efficient plan for brick procurement, but not
necessarily creative in exploring the space of all possible LEGO designs.

From the perspective of quantum physics, engineering is phenomena-based.
We humans wish to describe reality and observe nature. We discover objects
and effects in nature, and afterwards we may
then discover that they
can be useful in performing some useful task. An example is
the discovery of quantum tunnelling and its subsequent use to build a
field-effect transistor (FET) for digital electronics. This perspective asks the
questions: what is available to us, and what could we do it? It is like a child
going into a LEGO store with no pre-conceived plan, investigating all the
parts, dreaming up ways to use them, or even delving into custom brick design
as a more fundamental level. This will generate completely new
designs and models, but it is much slower than brick-shopping.

These perspectives are intertwined and serve complementary purposes,
and we often alternate
between the two of them. In our transistor example, we probably would not
have thought to build a FET if we had not already needed fast electrical
switching, for which vacuum tubes were proving inadequate. Moreover, we can
go beyond using quantum physics to simulate a classical device. There is
a remarkable example
of a new class of algorithms designed using the building blocks of
quantum physics. The first instance of such an algorithm solves the problem
of factoring.

% II. Factoring
The problem of factorization, or factoring,
is simple to state and difficult to solve.
Given an integer $m$, return the integers $\{p_1, \ldots, p_k\}$ whose
product of their powers is $m$: $m = p_1^{\alpha_1}\cdots p_k^{\alpha_k}$.
It is a beautiful theoretical problem posed by Carl Friedrich Gauss in 1801;
although he is considered by many to be one of history's greatest mathematicians,
he was still unable to factor arbitrarily large numbers. 
His feelings on the subject can be paraphrased as follows:
Factoring was the most important problem in the theory of
numbers, itself the crown jewel of mathematics, itself
the queen of the sciences.\footnote{\emph{Gauss zum Ged\"achtniss} (1856) by Wolfgang Sartorius von Waltershausen}
However, all was not lost.
It is a quirk of cryptography that the hardness of problems can be used
to secure communication. Therefore, factoring also has immense practical
usefulness as the basis of the RSA cryptosystem, the most widely-used
cryptosystem around the world and on the internet.

The somewhat sinister political importance of factoring to
governments is the ability to gather intelligence on each other and on their
citizens. The importance to citizens is the ability to choose security
parameters so that their correspondence is protected from each other and
their governments for a limited time.
The organizers of the popular democratic uprisings in the Middle East known
as the Arab Spring in 2011 definitely benefitted from using a cryptosystem that was
secure against compromise given the resources of the regimes that they toppled.
A government able to hack the organizers' Twitter accounts or Facebook pages
could misdirect protest actions or identify activists for imprisonment.
Similarly, being able to accurately estimate the resources needed to factor
an RSA key of a certain size may help protect future citizens against
governments with a quantum computers, as governments will be the first
organizations interested in and able to afford such a device.

How can we reason about the computational resources needed to solve the
problem of factoring an $n$-bit number?
Before 1994, the best known algorithm was the
number field sieve (citation needed) which had a running time of
$exp(\sqrt[3]{log n})$. This can be considered a pre-quantum (classical)
algorithm in a pre-quantum era. Its sub-exponential (but still super-polynomial)
running time is still considered intractable.
This conjectured hardness is why users of the RSA cryptosystem above have faith
that their information is resistant to compromise, while at the same time
giving hope to mathematicians of some future improvement.

In 1994, Peter Shor fulfilled this hope in a landmark
discovery that using the more general computational power of quantum physics,
factoring could be performed in running time $O(n^3\log n \log\log n)$\cite{Shor1994}. 
This kickstarted the entire field of quantum computing; united many
disciplines including computer science, physics, engineering, and
mathematics; and provided the foundation for this dissertation.
The computational model in which this remarkable feat was achieved
made some assumptions about quantum computing devices, which are now
standard in the field.
%However, many remain skeptical of this achievement due to the perceived
%plausibility of these assumptions, which is why RSA still remains the
%most widely-used cryptosystem in the world.\footnote{Despite recommendations
%from the NSA to switch to ECC, itself susceptible to quantum attacks (citation needed).}
These assumptions include robust error-correction (to preserve fragile
quantum information against environmental noise) and high-fidelity
quantum gates (to exert reliable classical control).
Whether these assumptions can be
realized is an active area of experimental physics research, and is not the
topic of this current work. However, given that these assumptions will
eventually be realized, how can we optimize Shor's factoring algorithm?
Quantum computation is now a rapidly maturing field, and we cannot provide
an adequate background in this dissertation. The interested reader is referred
to the standard textbook in the field by Nielsen and Chuang \cite{Nielsen2000}.
Building upon this background, we can provide an introduction into the subfield
of quantum architecture with the current work.

% III. Quantum circuits and quantum architecture
First, we will introduce a quantum circuit.
Analogous to the classical circuit model, a quantum circuit consists of
a network, or graph, of gates operating on quantum bits (qubits). One can
arrange this graph such that directed edges represent qubits and
nodes represent reversible quantum gates with in-degree equal to out-degree.
See Figure \ref{fig:intro-qcirc} for a representation of a quantum circuit
as a special class of directed
acyclic graphs. All flows travel from the input state of the qubits at the
first timestep (on the far
left) to the final output state on the same qubits in the last timestep
(on the far right). This induces a natural definition of circuit depth,
size, and width, which we will make more formal later.

\begin{figure}
\begin{array}{c}
\text{Size} = 15 \\
\Qcircuit @C=0.5em @R=.5em { 
	& \gate{ } & \qw & \ctrl{1} & \gate{ } & \qw & \qw      & \ctrl{1} & \qw \\ 
	& \gate{ } & \qw & \gate{ } & \ctrl{2} & \qw & \gate{K} & \targfix & \qw \\
	& \gate{ } & \qw & \gate{ } & \qw      & \qw & \gate{H} & \qw      & \qw \\
	& \gate{ } & \qw & \ctrl{1} & \gate{ } & \qw & \gate{H} & \qw      & \qw \\
	& \gate{ } & \qw & \gate{ } & \gate{ } & \qw & \qw      & \qw      & \qw
	\gategroup{1}{2}{5}{9}{.7em}{--}
}\\
\xymatrix {
  & & \text{Depth} = 5 \ar[ll] \ar[rr] & & \\
 }
\end{array}

\caption{A quantum circuit as a directed acyclic graph.}
\label{fig:intro-qcirc}
\end{figure}

A quantum architecture is a quantum circuit with constraints. For this
dissertation, we will concern ourselves with two main constraints:
(1) the layout graph of qubits which constrain two-qubit gate interactions and
(2) the decomposition of all single-qubit gates to a discrete, finite,
set of gates which are universal combined with a fixed two-qubit gate.
Quantum architecture provides an intermediate layer in between working
with actual hardware (the physical technology in a laboratory) and the
abstract theory (algorithms divorced of all physical constraints).
As quantum architects, we can optimize algorithms in a general way which
is relevant to many kinds of laboratory experiments. At the same time,
we can take into account the more complicated nature of physical devices
to provide more pessimistic, accurate, and numerical resource estimates than more
high-level theoretical results.

The role of quantum architecture is two-fold. First,
to quantify and improve the resources---such as circuit depth,
size, and width mentioned above---required to run a quantum
algorithm on a system of varying degrees of realism. Second, to provide
tradeoffs between circuit resources, or conceptual knobs used to tune a
particular implementation. This allows experimentalists to configure a
quantum architecture to meet their needs in a laboratory while allowing
theorists to improve tradeoffs or design new algorithms which take
physical constraints into account.


% Point 
We are now ready to state the thesis that we investigate in this dissertation.

\begin{quote}
Studying the depth of quantum architectures can help us design quantum computers that can solve
human problems with a human lifetime.
\end{quote}

Here, or somewhere else, we must distinguish between quantum architecture as a field and quantum architecture as a
circuit with hardware constraints. Actually, the two definitions are pretty close.

We also need to provide preliminary definitions of a quantum bit.

This deserves some unpacking. A quantum architecture is a circuit designed to implement a particular algorithm given certain
hardware constraints. The depth of a quantum architecture is the amount of time it takes for the circuit to complete.
Given a particular input $x$, how long (in human clock time) does it take to return the output of some function $f(x)$?
Quantum computers are interesting because for some algorithms, namely factoring, the depth of a quantum circuit is
less than the depth of any known classical circuit. However, this depth may correspond to a running time which is
still much longer than a single human lifetime. In particular this author, who is 32 years old at the time of this
writing, is interested in techniques to optimize the successful completion of a problem within his own lifetime.

Therefore, one meaning of ``studying the depth'' means to \emph{decrease} depth. The main technique for decreasing
depth is to parallelize a quantum circuit, that is, to divide and duplicate information to allow them to be processed
in a parallel way, using multiple classical controllers. However, parallelizing can increase other circuit resources,
usually due to this information duplication. This introduces a time-space tradeoff. In practical reality, we cannot
decrease depth at the expense of other circuit resources, such as the amount of information in a form suitable for
operation over time.

Decreasing depth can increase circuit size (the total number of operations to be performed) and circuit width
(the total amount of quantum information that can exist at any one time). This increase may again far
outstrip human resources, and delay completion of a quantum computer's construction, let alone its
successful execution, until well after that author's death. Therefore, the useful study of depth to help
quantum computers solve human problems must also limit this time-space tradeoff overhead. It should also
involve defining the exact tradeoff that is useful to minimize.

We have outlined the problems of designing such a low-depth quantum architecture so far in a very general,
abstract form. In this thesis, we will make these concepts more concrete by describing a formalism for
quantum circuits and quantum computation that is well-known to the research community.
Building on this formalism, we will propose a new architectural model and present our solutions to factoring in them.


Section \ref{sec:intro-arch} will discuss architectural models. We will present the definition of an
architectural models and give examples of them from the
current literature, with names like \textsf{AC} and \textsf{NTC}. We will discuss the idea of architectural
sub-models, with names like \textsf{2D CCNTC}. Finally,
we will describe our contribution of a new architectural model, called
\textsf{2D CCNTCM}, which is a weakly 2D,
hybrid architecture that introduces the idea of allowing both short-range
(local) interactions and long-range
interactions, and providing a configurable tradeoff in between them.

In order to build long-range interactions out of short-range interactions,
Section \ref{sec:intro-cdc} 
gives a thorough background on recent constant-depth communication techniques. These include three
basic constructs for teleporting, copying, and uncopying quantum information across an unbounded
lattice.

All operations on a quantum architecture can be reduced to single-qubit operations and two-qubit operations.
The latter operations are what we have been calling ``interactions,'' and we have studied how to overcome
constraints on nearest-neighbor interactions using the constant-depth communication techniques previously.

Ultimately, however, the limits to depth of a quantum architecture are
dependent on the single-qubit operations. 

The remainder of this chapter is devoted to the necessary background needed
to understand quantum architecture in general and factoring architectures
in particular. Quantum architecture has potentially many concerns, but we
focus on two of them: interaction distance and fault-tolerant quantum compiling.
Without loss of generality, the gates of a quantum circuit can be reduced to
single-qubit and two-qubit gates, called a circuit basis.
In fact, we can use a fixed two-qubit gate
and combine it with arbitrary single-qubit gates.
Section \ref{sec:intro-basis} provides the necessary background for
understanding circuit bases and single-qubit gates,
which do not impose any constraints on
which qubits can interact with each other. Section \ref{sec:intro-arch}
defines quantum architectures and architectural models in more detail using
both a fixed, finite basis in both single-qubit and two-qubit operations
and imposing connectivity constraints on the two-qubit operations.
Section \ref{sec:intro-cdc} introduces recent constant-depth techniques
for simulating long-distance interactions with nearest-neighbor
interactions, using linear size and width. Constant-depth communication
introduces a tradeoff between depth and the other two resources. Specifically,
reducing depth raises the possibility that more circuit resources are
devoted to communication than to computation. This tradeoff is examined 
in more detail in Section \ref{sec:intro-modules}
by introducing the notion of a module and hybrid nearest-neighbor
quantum architectures.

After this chapter, the remainder of this dissertation is organized to
examine and expound upon the thesis statement above. In the first two
chapters, we focus on improving the depth of factoring architectures
in our hybrid 2D nearest-neighbor model using parallelization and
constant-depth communication. In contrast to nearest-neighbor factoring
architectures, we will refer to our results as hybrid factoring architectures,
or simply hybrid factoring.
Chapter \ref{chap:factor-polylog} presents our first main result:
hybrid factoring in polylogarithmic depth.
Chapter \ref{chap:factor-sublog} presents a further exponential improvement
in our second main result: hybrid factoring in sublogarithmic depth.

At this point, we have run into a seemingly fundamental limit on improving
hybrid factoring to be constant depth: the compilation of arbitrary
single-qubit gates to a fault-tolerant basis. We examine this
\emph{quantum compiling} limit in Chapter \ref{chap:qcompile}. First we
present a background of quantum compiling and a survey of the current
state-of-the-art in related literature to provide context. Then, in our
third main result, we improve
upon the Kitaev-Shen-Vyalyi procedure for generating quantum Fourier states,
a vital subproblem for single-qubit quantum compiling.

Finally, we change directions and question the approach of minimizing depth
at all costs in the first two chapters. We argue that the most relevant
tradeoff to consider along with decreasing depth is the increasing amount
of error-correcting effort needed to maintain a useful quantum state for
computation. Therefore we introduce a new resource called
\emph{circuit coherence} in Chapter \ref{chap:coherence}. In this chapter,
we explore circuit coherence and its relationship to other circuit resources
in the hybrid model as well as other models of computation. In our final
main result, we show the connection between the circuit coherence for a
layered quantum circuit and a reversible pebble game, for which existing
time-space tradeoffs are known.

Finally we conclude our dissertation in Chapter \ref{chap:conclude} with a
summary of our main results along with related open conjectures and
directions for future research.

\input{intro/intro-basis.tex}

\input{intro/intro-arch.tex}

\input{intro/intro-cdc.tex}

\input{intro/intro-modules.tex}

% Old leftover stuff
So now we have discovered that quantum computing is theoretically interesting
because it can solve an important human problem. But can quantum computers be
actually built?

Quantum architecture is the intermediate layer between algorithms and hardware.
It is the design of how quantum bits and their allowed interactions in order
to solve these algorithms efficiently on realistic models of quantum hardware.
It aims to minimize circuit resources of interest, namely depth, size, and
width. In analogy to classical circuits, the depth is the running time of an
algorithm allowing parallelization, the size is the number of operations
required over all parallel processors, and the width is the number of
(quantum) bits required over all parallel processors.

In this dissertation, we study quantum architecture in the context of
optimizing Shor's factoring algorithm on nearest-neighbor architectures with
realistic constraints. It is hoped that lessons learned in this special-case
can contribute to the general community general principles which can be used
to generalize other quantum algorithms. We posit that the larger overall goal
of quantum architecture should be the design of a general-purpose quantum
processor, one which can execute any quantum algorithm with emphasis on being
able to perform a core group of operations efficiently. This core group of
operations is defined by the instruction set. Once quantum architecture has
progressed to this point, we can leverage the remarkable strides in normal
digital architecture over the past 80 years.

As well, we may be able to use the insights of quantum architecture to build
quantum processors which can surpass digital architectures in the solution of
large-scale problems over exponentially-sized solution spaces.
