\chapter{Quantum Circuit Coherence}
\label{chap:coherence}

In the first two chapters, we presented low-depth nearest-neighbor
architectures for factoring an $n$-bit number. We improved the depth first to
be sublinear and then sublogarithmic, but at a polynomial increase in size
and width. This represents a time-space tradeoff which can be upper-bounded
by the product of the circuit depth and circuit width. Although the title of
this dissertation indicates that it is beneficial to decrease depth, in
experimental implementations, we may be constrained by other real-world
resources.

Toward this end, we introduce a new circuit resource called \emph{coherence}
which quantifies the amount of error-correction that must be performed
to maintain a coherent quantum computing state. This can be analogous to
the amount of classical controller time or electrical power that a
quantum computing experiment consumes while running an algorithm. We
define our new resource in Section \ref{sec:cohere-def}, discuss its
relationship to other circuit resources on \textsf{2D CCNTCM}, and
calculate coherence for some simple examples to illustrate how it
captures the notion of parallelizability for a circuit.

Decreasing circuit coherence can increase circuit depth or size, which introduces
a new tradeoff for quantifying circuit parallelism.
Therefore, we compare it to other (time-space) tradeoffs that are common in
the literature, including measurement-based quantum computer (MBQC) in
Section \ref{sec:cohere-mbqc} and the reversible pebble game in
Section \ref{sec:cohere-tradeoff}. In both cases, we relate these
other tradeoffs to circuit coherence.

Finally, in Section \ref{sec:cohere-factor}, we calculate the circuit coherence
for an arithmetic building block useful in factoring: modular multiple addition.
We present fascinating conjectures about ways to further decrease
either depth or coherence for factoring architectures. Finally, we conclude
by presenting configurable-depth factoring circuits
that a future experimental architect can use to make the appropriate choice.

%Finally, having exhausted our fascination with factoring,
%we apply our low-depth techniques (compiling and circuit coherence)
%to a new quantum algorithm: hamiltonian
%simulation. In Section \ref{cohere-hs-bg}, we discuss the background of
%this problem, and in Section \ref{cohere-hs-calc} we parallelize one
%aspect of it, that of decreasing the depth of simulating
%a $1$-sparse Hamiltonian matrix.

\input{coherence/coherence-def.tex}

\input{coherence/coherence-mbqc.tex}

\input{coherence/coherence-tradeoff.tex}

\input{coherence/coherence-factor.tex}

%\input{coherence/coherence-hs-bg.tex}

%\input{coherence/coherence-hs-calc.tex}